### 2024-05-30 Paper/paper.tex
Updates from Overleaf

    \begin{abstract}
		{\small
        In many industries, not-for-profits (NFP) are the dominant firm type. Understanding the objectives of these firms is important, but NFP firms have underlying goals that make understanding their objectives difficult. For hospitals in particular, there is mixed evidence on whether NFP hospitals differ meaningfully from for-profit (FP) hospitals. In this paper, I investigate whether leadership team composition of NFP hospitals affects how similarly they behave to for-profits. I use a novel data set on NFP hospital executives in the US (2009-2015) to investigate whether the presence of clinically trained executives affects hospital behavior. In particular, I leverage pay-for-performance initiatives in the US enacted in 2012, and measure the  readmission and mortality rate responses between FP, NFP, and NFP with and without clinically trained executives. I find that NFPs without clinically trained executives behave more like FPs than other NFPs. This is consistent with theoretical predictions under the assumption that NFP hospitals with clinically trained executives value societal outcomes more than those without clinically trained executives.
		} 
	\end{abstract}

@@

    \subsection{Not-for-Profit Hospital Leadership}

    Not-for-profit firms differ from for-profits mainly in that for-profit firms distribute revenue to stakeholders, while not-for-profits reinvest profit back into the firm. Another important distinction is that NFPs are, in general, driven by a mission to better society, and therefore receive tax benefits.

    Hospitals are governed by a board of directors, whose role is to set broad goals and strategies, and provide general oversight. The board selects executives, the highest level of management of a firm, to carry out day-to-day operations. An executive team usually consists of at least a CEO and CFO, but every firm is different in exactly how they structure their executive team. Many hospital executives specialize in health care settings by getting degree(s) in health care management, or an MBA specific to health care. There are specific positions that are often filled by someone with clinical training, such as a Chief Medical Officer, and doctors can even find themselves in other top managerial positions as well, such as CEO or president. While some doctors earn additional degrees before stepping into an executive role, this is not a necessary condition to becoming a physician executive.

    Anecdotal evidence is mixed on the benefits of hiring physician executives. On one hand, physicians bring a unique combination of clinical expertise and being comfortable with administration, and they can bring insights that improve hospital outcomes (\cite{Stajduhar_2023}, \cite{Ahmed_2022}). Alternatively, physicians may not always be adequately trained to enter into high level management positions, and therefore are not equipped to lead a firm (\cite{HarvardBusinessReview2018}).

    All tax-exempt organizations in the US are required to file a Form 990 with Internal Revenue Services (IRS) each year. There are different types of forms, but any organization grossing over \$200,000 must file the most extensive Form 990. Sections of this form include a statement of revenue, statement of functional expenses, a balance sheet, and, as used in this project, a list of all key employees, executives, and board members. Each firm is required to report the name and title, average hours per week, position, and compensation of their board members and executive leaders.

@@

    Two programs were passed as part of the Affordable Care Act that focused on pay-for-performance incentives for hospitals, the Hospital Readmissions Reduction Program (HRRP), and the Hospital Value Based Purchasing Program (HVBP). HRRP focused on penalizing hospitals with poor quality in terms of readmission rates, and HVBP focused on rewarding hospitals with high quality or show improvement in quality.

    In October 2011, the Center for Medicare and Medicaid Services (CMS) released a set of rules under HRRP that mandated penalties for hospitals with above average readmission rates. A readmission is when a patient returns to the hospital within 30 days of being discharged from a previous stay; avoidable readmissions are a bad outcome for patients and increase health care spending. The goal of HRRP is to lower readmissions through better care coordination, less initial stay complications, and better post-care instructions. Beginning in October 2012, hospitals with higher readmission rates than the national average in pneumonia, heart failure, or AMI (after adjusting for demographic characteristics) receive a fixed lower reimbursement rate for all Medicare patients seen in their hospital. In 2015, CMS also included chronic obstructive pulmonary disease, coronary artery bypass graft surgery, and elective primary total hip arthroplasty and/or total knee arthroplasty as conditions which go into the penalty calculation (\cite{CMS}).

    Penalized hospitals pay penalties in the form of a fixed rate for every Medicare patient regardless of the condition. The rate began at 1\% and increased to 3\% in 2015. Further, CMS does not distinguish a necessary readmission from an avoidable readmission; any repeat hospital visit is included in the penalty calculation. Excess readmission rates are calculated using a rolling look-back period of 3 years to determine whether the hospital is penalized. Therefore, hospitals had incentive to react immediately once details of the program were announced in October of 2011. These penalties are not insignificant; penalized hospitals paid, on average, 4-5\% of revenue.

    The HVBP Program instead rewards hospitals with high quality or improvement in quality. Specifically, CMS deducts Medicare payments by 2\% from all eligible hospitals, collects this sum, and divides it among the rewarded hospitals. Several quality and cost measures surrounding safety, efficiency, cost reductions, clinical outcomes, and community engagement, are combined to create a single score metric for each hospital. Hospitals are then compared to the average and are rewarded for being above average quality and for showing improvement (\cite{CMS_2023}).

@@

    Hospital executive teams are understudied partly because of the inaccessibility of information on them. I construct a novel data set of not-for-profit hospitals in the US that contains names, titles, and positions of all board members and executives tied to the hospital from 2009-2015. I gather this data from publicly available Tax Form 990s, which every sufficiently large not-for-profit files each year in the US. To my knowledge, this is the first large scale gathering of executive names from these forms.\footnote{\citeauthor{brickley2010board} (\citeyear{brickley2010board}) collects compensation information for a select number of hospitals.}

    The historical tax forms for all not-for-profits are housed by ProPublica.\footnote{https://projects.propublica.org/not-for-profits/} I use the NonProfit Explorer API to extract Employee Identification Numbers (EIN) of NFPs classified as a hospital according to their National Taxonomy of Exempt Entities (NTEE) code. After filtering out associations and specialty hospitals, there are approximately 3,000 NFP EINs. Importantly, linking EINs to other hospital information is crucial for the analysis. I rely on matching by location and name of hospital in the tax forms and AHA data. I confidently match 1200 EINs to an AHA identification number based on exact name matches within the same state. I compare the matched hospitals to non-matched hospitals in Appendix \ref{app:matched}.

    For each of the 1200 EINs, I then extract web URLs to Form 990 PDFs in each year. I download these locally and use optical character recognition (OCR) text extraction methods to scrape the relevant section of the PDF, which is titled ``Officers, Directors, Trustees, Key Employees, and Highest Compensated Employees". Finally, I use string cleaning methods to identify names and positions for each EIN, year. I also identify titles indicating clinical training: MD, Dr., or DO. Using OCR extraction can be unreliable in some instances if the text is handwritten or small. Therefore, I manually record names, positions, and titles for observations that are missing after the initial extraction. This yields 852 not-for-profit hospitals that are matched to AHA data, and are not missing text information.

    Due to the lack of consistency in the organization of executive teams, I focus mainly on executive teams as a whole instead of specific positions. I drop anyone who is solely a board member, and focus on those who have executive in their title or are labeled as a president or vice president of a department. I create various hospital-level characterizations based on their executive team, such as the existence of a clinically trained executive, the number of clinically trained executives, the number of total executives, and whether the hospital has a Chief Medical Officer. Additionally, I examine changes in executive teams to determine whether the hospital changes its propensity to hire a clinically trained executive. Summary statistics for these variables are shown in Table \ref{leader_sumstats}. On average, NFP hospitals in my sample have 4 executives and .4 clinically trained executives. Sixty-one percent of the sample never changes propensity to hire a clinically trained executive, and 33\% of NFPs in the sample have a Chief Medical Officer.

@@

    I merge the created sample of not-for-profits with all for-profits in the AHA survey. The outcomes I consider are readmission and mortality rates, which are directly targeted in the pay-for-performance policies, and are found in the CMS Hospital Compare data. I combine readmission and mortality rates for pneumonia, AMI, and heart failure (the relevant HRRP conditions) into a weighted average based on the number of patients in each condition. Additionally, I include several relevant hospital characteristics. Number of beds, whether the hospital is an academic medical center, is physician owned, or is system affiliated, all come from the AHA survey. Uncompensated care comes from HCRIS, and case mix index, which measures patient complexity, comes from CMS. Whether the hospital was penalized under HRRP or received payments under HVBP is found in the HCRIS data. Finally, I limit to years 2010-2014.

@@

    I present a table of means for each hospital type (for-profit, in-sample not-for-profits, and not-for-profits with and without a clinically trained executive) in Table \ref{tab:sumstats_samples}. Approximately 850 NFPs have leadership information, with 384 of them hiring an MD executive at some point in the sample. I compare these NFPs with the 740 for-profits from the AHA Survey. For-profits and NFPs are roughly the same bed size, but NFPs see more patients in each of the three relevant conditions. For-profits are more likely to be penalized, but also more likely to receive HVBP incentives (potentially for improvement in quality). For-profits have higher readmission rates, and higher mortality rates in all categories except heart failure. There are also key differences in not-for-profits with and without clinically trained executives. Those with MDs are larger, more likely to be penalized, have higher uncompensated care, and more complex patients. I discuss the implications of these differences for identification in Section \ref{sec:model}.

@@

    I model hospital behavior in two simplified scenarios: where quality does not directly affect profit, and where it does. Pay-for-performance incentives essentially incorporate performance directly into a hospital's profit function. Here, I model performance as quality of care, which is inversely related to readmission and mortality rates. While the literature speculates about the true form of NFP objective functions, we can generally think of hospitals choosing some linear combination of profit and societal benefit. Intuitively, a hospital that places only weight on profit (for-profit type) will only care about quality when it enters into the profit function, and therefore will respond more drastically to pay-for-performance than a hospital who already cared about quality through societal benefit before the incentives. Therefore, hospital response to pay-for-performance incentives reveals something about their underlying preference on profit vs. societal outcomes.

    To formalize this intuition, I specify the hospital objective function as a weighted average of profit and societal (or patient) outcomes and derive characteristics of the chosen quality. For simplicity, I abstract away from quality having a direct effect of quantity and price. Without any financial incentives to increase quality, hospital quality is decided by:

@@

    \noindent marginal benefit equals marginal cost of increasing quality. Solving for $u'(\theta)$ and differentiating with respect to $\alpha$ yields

@@

    Thus,

@@

    Therefore, in a world with financial incentives on quality, quality is increasing with more weight placed on profit. I combine the results found in each scenario into one response function that depends on $\alpha$, where $\theta_2$ is quality under pay-for-performance incentives and $\theta_1$ is quality with no financial incentive to quality. That is,

@@

    Hence, this model predicts that change in quality depends on how much weight the hospital places on profit vs. societal benefit. Particularly, hospitals with more weight on profit respond more than hospitals with more weight placed on societal benefit. In reality, $\alpha$ is unknown, except in the case of for-profits, who have revealed that they have a high $\alpha$ by their ownership status. However, not-for-profit behavior on average does not bring much clarity about their objectives. An observable and economically relevant characteristic of a hospital that could affect $\alpha$ is composition of their leadership team. Thus, I compare the response of for-profits and not-for-profits with different leadership teams to bring clarity on how this might point to underlying $\alpha$.

@@

    In light of the theoretical predictions, I first test empirically whether NFPs respond to program incentives differently than for-profits. I estimate the following regression equation:

    \begin{equation}
    \label{eq:forprofit}
    y_{ht} = \beta (\text{For-Profit}_h \times \text{Post Programs}_t) + \gamma_{h} + \delta_t + \epsilon_{ht}
    \end{equation}

    \noindent where $y_{ht}$ is one of the outcome variables discussed in Section \ref{sec:data}, $\beta$ is the coefficient of interest capturing the effect of being for-profit after the programs were introduced, and $\gamma_h$ and $\delta_t$ are hospital and time fixed effects, respectively. For each outcome, I estimate three models with differing comparison groups: all NFPs, NFPs with a clinically trained executive, and NFPs without a clinically trained executive. I also directly estimate the difference in response of NFPs with and without a clinically trained executive:

    \begin{equation}
    \label{eq:clinical}
    y_{ht} = \beta (\text{No MD Exec}_h \times \text{Post Programs}_t) + \gamma_{h} + \delta_t + \epsilon_{ht}
    \end{equation}

@@

    Additionally, under certain assumptions that I discuss in Section \ref{sec:identification}, using executive team changes can be advantageous in better understanding underlying hospital motives. There are two mechanisms to having a clinically trained executives affecting outcomes: first, a clinically trained executive could be a signal of the underlying weight placed on profit, $\alpha$. Second, hiring a clinically trained executive who carries out day-to-day operations in a distinct way effectively changes $\alpha$. Thus, I use two types of variation in clinically trained executives to disentangle the two: a variable for whether the hospital ever hires a clinically trained executive, and a variable for whether the hospital has a clinically trained executive in 2012 when the programs took place. I use the same empirical specification with these two variables as the independent variation in executive teams, and limit the sample to the relevant comparison group:

    \begin{equation}
    \label{eq:decomp}
    y_{ht} = \beta (\text{Never/No 2012 Exec}_h \times \text{Post Programs}_t) + \gamma_{h} + \delta_t + \epsilon_{ht}
    \end{equation}

@@

    To identify the causal effect of hospital type on hospital program response, I rely on several assumptions. First, I assume that, absent the program enactments, different hospital types would have had parallel trends in outcomes. While I see no evidence of pre-trends, Table \ref{tab:sumstats_samples} suggests that hospitals with a clinically trained executive are different than other hospitals along multiple characteristics. Therefore, there may be concern that these differences between hospital types could be driving the results. Therefore, my main specification uses the synthetic difference-in-differences method established in \citeauthor{arkhangelsky2021synthetic} (\citeyear{arkhangelsky2021synthetic}). This method creates unit weights that more heavily count comparison hospitals that are similar to treatment hospitals prior to the programs. It also weighs time periods that balance pre-program and post-program for the comparison hospitals. Then, a simple two way fixed effects estimator is used with these weights, yielding a more robust average treatment effect that relies less on a strong parallel trends assumption (\cite{arkhangelsky2021synthetic}). I present the synthetic diff-in-diff estimates as the main results, and they are largely similar to the standard two way fixed effects results in Appendix \ref{app:fullsample}. Further, while ultimately I employ a synthetic difference-in-differences method, the conceptual identification is the same as a standard difference-in-differences.

@@

     I assess the validity of this assumption in two ways. First, I analyze whether changes in clinically trained executives occur more often in time periods of program enactment. Second, I analyze whether hospitals who end up getting penalized or receiving payments are more likely to make changes in clinical executive members than those who are not. That is, are hospitals that expect to be affected by the programs more likely to hire or fire clinically trained executives? I consider changes along having any clinically trained executive as well as changes in the number of clinically trained executives. I estimate the following regression equations:

@@

    \begin{equation}\label{eq:change2}
    \text{change}_{ht} = \beta(\text{program\_exposed}_{h} \times \text{Post Programs}_t)+ \alpha_h + \delta_t + \epsilon_{ht}.
    \end{equation}

    The variable $\text{program\_exposed}_{h}$ is an indicator for whether the hospital eventually was penalized under HRRP or received payments under HVBP. The estimates from this analysis are presented in Table \ref{tab:change_analysis}. Changes in any clinically trained executive are less likely to occur in 2012-2014 relative to 2010, indicating that executive teams are actually becoming more stable over time along the dimension of hiring clinically trained executives. Further, hospitals who end up being penalized through HRRP or receiving payments through HVBP are not more likely to change their propensity to hire any clinically trained executive. Results are similar when considering changes in the number of clinically trained executives. Thus, it seems unlikely that endogenous team formation is biasing the estimates.

@@

     The estimated difference in readmission rate trends for all hospital types in response to the programs, along with ATE and standard errors, are shown in Figure \ref{fig:read_synth_plot}. I first discuss the differential program response of for-profit firms compared to the different NFP sub-samples (Equation \ref{eq:forprofit}), shown in panels \ref{fig:read_synth_plota}-\ref{fig:read_synth_plotc}.

     Panel \ref{fig:read_synth_plota} shows that on average, there is no difference in how for-profits and not-for-profits change their readmission rates in response to the incentive changes. However, when NFPs are subset into whether they have a clinical executive or not, panel \ref{fig:read_synth_plotb} shows that there is a differential response to the program between for-profits and NFPs with a clinical executive. In particular, for-profits decrease readmission rates by .3 more than NFPs with a clinical executive, an additional 3 patient readmission decrease relative to the average number of patients. While numerically small, this difference in response is non-existent between for-profits and NFPs without a clinical executive (panel \ref{fig:read_synth_plotc}), indicating some, if minor, difference in objectives resulting from having a clinical executive.

@@

    Next, in panel \ref{fig:read_synth_plotd}, I present estimates of Equation \ref{eq:clinical}, directly comparing NFPs with and without a clinical executive. NFPs without a clinically trained executive lowered readmission rates by .24 ppts more than NFPs with a clinically trained executive.

>>>>

    \begin{abstract}
		{\small
        In many industries, not-for-profits (NFP) are the dominant firm type. Understanding the objectives of these firms is important, but NFP firms have underlying goals that make understanding their objectives difficult. For hospitals in particular, there is mixed evidence on whether NFP hospitals differ meaningfully from for-profit (FP) hospitals. In this paper, I investigate whether leadership team composition of NFP hospitals affects how similarly they behave to for-profits. I use a novel data set on NFP hospital executives in the US (2010-2014) to investigate whether the presence of clinically trained executives affects hospital behavior. In particular, I leverage pay-for-performance initiatives in the US enacted in 2012, and measure the  readmission and mortality rate responses between FP, NFP, and NFP with and without clinically trained executives. I find that NFPs without clinically trained executives behave more like FPs than other NFPs. This is consistent with theoretical predictions if NFP hospitals with clinically trained executives value societal outcomes more than those without clinically trained executives.
		} 
	\end{abstract}

@@

    \subsection{Not-for-Profit Hospital Executives}

    Not-for-profit firms in the US differ from for-profits mainly in that for-profit firms distribute revenue to stakeholders, while not-for-profits reinvest profit back into the firm. An important distinction between the two is that NFPs are, in general, driven by goals to better society, and therefore receive tax benefits from the government.

    Hospitals are governed by a board of directors, whose role is to set broad goals and strategies, and provide general oversight. The board selects executives, the highest level of management of a firm, to carry out day-to-day operations. An executive team usually consists of at least a Chief Executive Officer (CEO) and Chief Financial Officer (CFO), but there is variation in how firms organize executive teams. It is not uncommon for hospital executives to specialize in health care settings by earning a degree in health care management, or an MBA specific to health care. There are executive positions that are often filled by someone with clinical training, such as a Chief Medical Officer (CMO), and doctors can even find themselves in other top managerial positions as well, such as CEO or president. While some doctors earn additional degrees before stepping into an executive role, this is not a necessary condition to becoming a physician executive.

    Hiring physician executives can be controversial. On one hand, physicians bring a unique combination of clinical and administration expertise, and they have insights that can potentially improve hospital outcomes (\cite{Stajduhar_2023}, \cite{Ahmed_2022}). Alternatively, physicians may not always be adequately trained to enter into high level management positions, and therefore worsen hospital outcomes (\cite{HarvardBusinessReview2018}).

    All tax-exempt organizations in the US are required to file a Form 990 with Internal Revenue Services (IRS) each year. There are different types of forms, but any organization grossing over \$200,000 must file the most extensive Form 990. Sections of this form include a statement of revenue, statement of functional expenses, a balance sheet, and, as used in this project, a list of all key employees, executives, and board members. Each firm is required to report the name and title, average hours per week, position, and compensation of their board members and executives.

@@

    Two programs were passed as part of the Affordable Care Act that focused on pay-for-performance incentives for hospitals: the Hospital Readmissions Reduction Program (HRRP), and the Hospital Value Based Purchasing Program (HVBP). HRRP focused on penalizing hospitals with poor quality, and HVBP focused on rewarding hospitals with high quality or show improvement in quality.

    In October 2011, the Center for Medicare and Medicaid Services (CMS) released a set of rules under HRRP mandating penalties for hospitals with above average readmission rates. A readmission is when a patient returns to the hospital within 30 days of being discharged from a previous stay; avoidable readmissions are a bad outcome for patients and increase health care spending. The goal of HRRP is to lower readmissions through better care coordination, less initial stay complications, and better post-care instructions. Beginning in October 2012, hospitals with higher readmission rates than the national average in pneumonia, heart failure, or AMI (after adjusting for demographic characteristics) receive a fixed lower reimbursement rate for all Medicare patients seen in their hospital. In 2015, CMS also included chronic obstructive pulmonary disease, coronary artery bypass graft surgery, and elective primary total hip arthroplasty and/or total knee arthroplasty as conditions which go into the penalty calculation (\cite{CMS}).

    Penalties are given in the form of a fixed rate reduction of 1-3\% for every Medicare patient regardless of the condition. Further, CMS does not distinguish a necessary readmission from an avoidable readmission; any repeat hospital visit is included in the penalty calculation. Excess readmission rates are calculated using a rolling look-back period of 3 years to determine whether the hospital is penalized. Therefore, hospitals had incentive to react immediately once details of the program were announced in October of 2011. These penalties are not insignificant; penalized hospitals paid, on average, 4-5\% of revenue.

    The HVBP Program instead rewards hospitals with high quality or significant improvement in quality. Specifically, CMS deducts Medicare payments by 2\% from all eligible hospitals, collects this sum, and divides it among the rewarded hospitals. Several quality and cost measures surrounding safety, efficiency, cost reductions, clinical outcomes, and community engagement are combined to create a single score metric for each hospital. Hospitals are then compared to the average and are rewarded for being above average quality or for showing improvement (\cite{CMS_2023}).

@@

    Hospital executive teams are understudied partly due to the inaccessibility of granular data. I construct a novel data set of not-for-profit hospitals in the US that contains names, titles, and positions of all board members and executives tied to the hospital from 2009-2015. I gather this data from publicly available Tax Form 990s, which every sufficiently large not-for-profit files each year in the US. To my knowledge, this is the first large scale gathering of executive names from these forms.\footnote{\citeauthor{brickley2010board} (\citeyear{brickley2010board}) collects compensation data for a select number of hospitals.}

    The historical tax forms for all not-for-profits are housed by ProPublica.\footnote{https://projects.propublica.org/not-for-profits/} I use the NonProfit Explorer API to extract Employee Identification Numbers (EIN) of NFPs with a hospital National Taxonomy of Exempt Entities (NTEE) code. After filtering out associations and specialty hospitals, there are approximately 3,000 NFP EINs. Linking these EINs to other hospital information is crucial for the analysis. I therefore rely on matching by location and name of hospital in both the tax forms and AHA data. I confidently match 1200 EINs to an AHA identification number based on exact name matches within the same state. I assess the observable differences between matched and non-matched hospitals in Appendix \ref{app:matched}, and conclude that the samples are overall similar apart from an under-sampling of hospitals that belong to systems.

    For each of the matched EINs, I extract web URLs to Form 990 PDFs for each year. I use an algorithm to download these locally and use optical character recognition (OCR) text extraction methods to scrape the relevant section of the PDF, which is titled ``Officers, Directors, Trustees, Key Employees, and Highest Compensated Employees". Finally, I use string cleaning methods to identify all names and positions under each EIN, year. With each name, I extract titles indicating clinical training: MD, Dr., or DO. Using OCR extraction can be unreliable in cases where the text on the form is handwritten or small. Therefore, for observations that are missing after the initial extraction, I manually record names, positions, and titles. This yields 852 not-for-profit hospitals that are matched to AHA data, and are not missing text information.

    For the purposes of this paper, I focus mainly on executive teams as a whole instead of specific positions. The main reason for this is because the organization of teams varies so drastically across hospitals that a specific position, even if it exists in two different hospitals, can be meaningfully different. Additionally, the more granular the information I extract, the more likely there will be a mistake in the OCR extraction. Thus, to preserve correctness of the independent variable of interest, I focus on whether there is any MD executive at the hospital.

    I drop anyone who is solely a board member, and focus on those who have executive in their title, or are labeled as a president or vice president of a department. I create various hospital-level characterizations based on their executive team, such as the existence of a clinically trained executive, the number of clinically trained executives, the number of total executives, and whether the hospital has a Chief Medical Officer. Additionally, I examine changes in propensity to hire a clinical executive. Summary statistics for these variables are shown in Table \ref{leader_sumstats}. On average, NFP hospitals in my sample have 4 executives and .4 clinically trained executives. Sixty-one percent of the sample never changes propensity to hire a clinically trained executive, and 33\% of NFPs in the sample have a Chief Medical Officer at some point.

@@

    I combine the created sample of not-for-profits with all for-profits in the AHA survey. The outcomes I consider are readmission and mortality rates, which are directly targeted in the pay-for-performance policies, and are found in the CMS Hospital Compare data. I combine rates for pneumonia, AMI, and heart failure (the relevant HRRP conditions) into a weighted average based on the number of patients in each condition. Additionally, I include several relevant hospital characteristics: number of beds, whether the hospital is an academic medical center, is physician owned, or is system affiliated, all come from the AHA survey. Case mix index, a measure of overall patient complexity, comes from CMS. Whether the hospital was penalized under HRRP or received payments under HVBP is found in the HCRIS data. Finally, I limit to years 2010-2014.

@@

    There are several samples of hospitals that are relevant for the analysis. First, there are all for-profits and all not-for-profits. Further, I subset the sample of NFPs into whether they have a clinical executive at some point in the sample. Finally, I characterize a NFP leadership team as being stable or not. A stable team means the hospital did not change their propensity to hire a clinical executive over time. I present a table of means for each of these sub-samples in Table \ref{tab:sumstats_samples}. Approximately 850 NFPs have leadership information, with 384 of them hiring an MD executive at some point in the sample. I compare these NFPs with the 708 for-profits from the AHA Survey. For-profits and NFPs are roughly the same bed size, but NFPs see more patients in each of the three relevant conditions. For-profits are more likely to be penalized, but also more likely to receive HVBP incentives (potentially for improvement in quality). For-profits have higher readmission rates, and higher mortality rates in all categories except heart failure. There are also key differences in not-for-profits with and without clinically trained executives. Those with MDs are larger, more likely to be penalized, have higher uncompensated care, and more complex patients. This holds for stable leadership teams as well, though the number of NFPs with a clinical executive drops significantly. I discuss the implications of these differences for identification in Section \ref{sec:model}.

@@

    I model hospital behavior under two simplified worlds: one where quality does not directly affect profit, and one where it does. Pay-for-performance incentives essentially incorporate performance directly into a hospital's profit function. I directly model performance as quality of care, which is inversely related to readmission and mortality rates. While the literature speculates about the true form of NFP objective functions, we can generally think of hospitals choosing some linear combination of profit and societal benefit, without knowing how much weight the hospital places on each. Intuitively, a hospital that places only weight on profit (for-profit type) will only care about quality when it enters into the profit function, and therefore will respond more drastically to pay-for-performance than a hospital who already cared about quality through societal benefit before the incentives. Therefore, hospital responses to pay-for-performance incentives reveal information about their underlying preference on profit vs. societal outcomes.

    To formalize this intuition, I specify the hospital objective function as a weighted average of profit and societal (or patient) outcomes and derive characteristics of the optimal quality. For simplicity, I abstract away from quality having a direct effect of quantity and price. Without any financial incentives to increase quality, hospital quality is decided by:

@@

    \noindent where marginal benefit equals marginal cost of increasing quality. Solving for $u'(\theta)$ and differentiating with respect to $\alpha$:

@@

    Therefore, again using the Implicit Function Theorem,

@@

    That is, all else equal, in a world with financial incentives on quality, quality is increasing with more weight placed on profit. I combine the results found in each scenario into one response function that depends on $\alpha$, where $\theta_2$ is quality under pay-for-performance incentives and $\theta_1$ is quality with no financial incentive to quality. That is,

@@

    Hence, this model predicts that change in quality depends on how much weight the hospital places on profit vs. societal benefit. Particularly, hospitals with more weight on profit respond more than hospitals with more weight placed on societal benefit. In reality, $\alpha$ is unknown, except in the case of for-profits, who have revealed that they have a high $\alpha$ by their ownership status. However, not-for-profit behavior on average does not bring much clarity about their objectives. An observable and economically relevant characteristic of a hospital that could affect $\alpha$ is composition of their executive leadership team. Thus, I compare the response of for-profits and not-for-profits with different leadership teams to bring clarity on how this might point to underlying $\alpha$.

@@

    In light of the theoretical predictions, I test empirically whether different hospital types respond to pay-for-performance incentives differently. I present here linear models with a two-way fixed effects specification. Ultimately, I employ synthetic difference-in-differences, which is estimating the same model with weights that more heavily factor similar trending control groups and time periods. However, for clarity, I present the unweighted specification:

    \begin{equation}
    \label{eq:forprofit}
    y_{ht} = \beta \text{ treat}_{ht} + \gamma_{h} + \delta_t + \epsilon_{ht},
    \end{equation}

    \noindent where $y_{ht}$ is one of the outcome variables discussed in Section \ref{sec:data}, and $\gamma_h$ and $\delta_t$ are hospital and time fixed effects, respectively. The independent variable of interest, treat$_{ht}$, captures the effect of being a certain hospital type after incentives changed in 2012. That is, I interact an indicator for hospital type with an indicator for year post 2012 for various hospital types. First, I define treat so that for-profit hospitals are being compared to subsets of not-for-profits (by limiting the sample to the relevant treatment and comparison group). Then, I limit to only not-for-profits and define treat as those without a clinical executive. Specifically, I define treat$_{ht}$ and limit the sample in four ways, shown in Figure \ref{fig:spec}.

@@

\begin{figure}[ht!]
    \begin{center}
    \caption{\label{fig:spec}Treated and Comparison Groups}
 \begin{tabular}{| m{18em} |}
 \hline
 Combination 1:\\ [0.5ex]
 \hline\hline 
 \vspace{2mm}
 Treat Group: \hspace{23.5mm} For-Profit \\
 \vspace{2mm} 
 Comparison Group: \hspace{11.5mm} Stable NFPs  \\
 [1ex]
 \hline
 \end{tabular}
\hfil   %<---
 \begin{tabular}{|m{18em}|}
 \hline
 Combination 2:\\ [0.5ex]
 \hline\hline
 \vspace{2mm}
 Treat Group: \hspace{13.5mm} For-Profit \\
 \vspace{2mm}
 Comparison Group: \hspace{1.5mm} Stable NFP w/out MD  \\
 [1ex]
 \hline
 \end{tabular}

\medskip

\vspace{2mm}

  \begin{tabular}{|m{18em}|}
 \hline
 Combination 3:\\ [0.5ex]
 \hline\hline
 \vspace{2mm}
 Treat Group: \hspace{17mm} For-Profit \\
 \vspace{2mm}
 Comparison Group: \hspace{5mm} Stable NFPs w/ MD  \\
 [1ex]
 \hline
 \end{tabular}
\hfil   %<---
  \begin{tabular}{|m{18em}|}
 \hline
 Combination 4:\\ [0.5ex]
 \hline\hline
 \vspace{2mm}
 Treat Group:  \hspace{13mm} Stable NFPs w/out MD \\
 \vspace{2mm}
 Comparison Group:  \hspace{1mm} Stable NFPs w/ MD  \\
 [1ex]
 \hline
 \end{tabular}
 \end{center}
 \end{figure}

    Additionally, under certain assumptions that I discuss in Section \ref{sec:identification}, using executive team changes can be advantageous in better understanding underlying hospital motives. There are two mechanisms to having a clinically trained executives affecting outcomes: first, a clinically trained executive could be a signal of the underlying weight placed on profit, $\alpha$. Second, hiring a clinically trained executive who carries out day-to-day operations in a distinct way effectively changes $\alpha$. Thus, I also define treat and limit the comparison group in such a way that decomposes the two. Without limiting to stable teams, I consider hospitals that ever hire a clinically trained executive vs. hospitals that have a clinically trained executive in 2012 when the programs took place. These variations of Equation \ref{eq:forprofit} are shown in Figure \ref{fig:decomp_spec}.

    \begin{figure}[ht!]
    \begin{center}
    \caption{\label{fig:decomp_spec}Decomposition Model Specification Details}

 \begin{tabular}{| m{18em} |}
 \hline
 Decomposition 1:\\ [0.5ex]
 \hline\hline 
 \vspace{2mm}
 Treat Group:  \hspace{15mm} NFP w/out MD \\
 \vspace{2mm}
 Comparison Group: \hspace{3mm} NFPs w/ MD  \\
 [1ex]
 \hline
 \end{tabular}
\hfil   %<---
 \begin{tabular}{|m{18em}|}
 \hline
 Decomposition 2:\\ [0.5ex]
 \hline\hline
 \vspace{2mm}
 Treat Group: \hspace{11mm} No MD Exec 2012 \\
 \vspace{2mm}
 Comparison Group:  NFPs w/ MD (not 2012)  \\
 [1ex]
 \hline
 \end{tabular}

\end{center}
 \end{figure}

@@

    To identify the causal effect of hospital type on hospital program response, I rely on several assumptions. First, I assume that, absent the program enactments, treat and comparison group types would have behaved similarly in outcomes. While I see no evidence of pre-trends, Table \ref{tab:sumstats_samples} suggests that hospitals with a clinically trained executive are different than other hospitals along multiple characteristics. There may be concern that these differences between hospital types could be driving the results. Therefore, my main specification uses the synthetic difference-in-differences method established in \citeauthor{arkhangelsky2021synthetic} (\citeyear{arkhangelsky2021synthetic}). This method creates unit weights that more heavily factor comparison hospitals that are similar to the treat group hospitals prior to the programs. It also weighs time periods that balance pre-programs and post-programs for the comparison hospitals. Then, a simple two way fixed effects estimator is used with these weights, yielding a more robust average treatment effect that relies less on a strong parallel trends assumption (\cite{arkhangelsky2021synthetic}). I present the synthetic diff-in-diff estimates as the main results, and they are largely similar to the standard two way fixed effects results with no weights, shown in Appendix \ref{app:fullsample}. Further, the conceptual identification is the same as a standard difference-in-differences.

@@

     I assess the validity of this assumption in two ways. First, I analyze whether changes in clinically trained executives occur more often in time periods of program enactment. Second, I analyze whether hospitals who end up getting penalized or receiving payments are more likely to make changes in clinical executive members than those who are not. That is, are hospitals that expect to be affected by the programs more likely to hire or fire clinically trained executives? I consider changes in having any clinically trained executive, as well as changes in the number of clinically trained executives. I estimate the following linear equations:

@@

    \begin{equation}\label{eq:change2}
    \text{change}_{ht} = \beta(\text{Program Exposed}_{h} \times \text{Post Programs}_t)+ \alpha_h + \delta_t + \epsilon_{ht}.
    \end{equation}

    The variable $\text{Program Exposed}_{h}$ is an indicator for whether the hospital eventually was penalized under HRRP or received payments under HVBP. The estimates from this analysis are presented in Table \ref{tab:change_analysis}. Changes in any clinically trained executive are less likely to occur in 2012-2014 relative to 2010, indicating that executive teams are actually becoming more stable over time along the dimension of hiring clinically trained executives. Further, hospitals who end up being penalized through HRRP or receiving payments through HVBP are not more likely to change their propensity to hire any clinically trained executive. Results are similar when considering changes in the number of clinically trained executives. Thus, it seems unlikely that endogenous team formation is biasing the estimates.

@@

     The estimated difference in readmission rate trends for all hospital types in response to the programs are shown in Figure \ref{fig:read_synth_plot}. Each panel corresponds to a particular definition of the treat group of hospitals and limitation of comparison group hospitals.

     Panel \ref{fig:read_synth_plota} compares readmission rate responses among for-profits and all NFPs, and shows that on average, there is no difference in how these types of hospitals change readmission rates in response to the incentive changes. However, when the comparison group is limited to NFPs with a clinical executive, panel \ref{fig:read_synth_plotb} shows that there is a differential response to the program between these hospital types. In particular, for-profits decrease readmission rates by .3 more than NFPs with a clinical executive, an additional 3 patient readmission decrease relative to the average number of patients. While numerically small, this difference in response is non-existent between for-profits and NFPs without a clinical executive (panel \ref{fig:read_synth_plotc}), indicating some, if minor, differences in objectives resulting from having a clinical executive.

@@

    Next, in panel \ref{fig:read_synth_plotd}, I directly compare NFPs with and without a clinical executive. NFPs without a clinically trained executive lowered readmission rates by .24 ppts more than NFPs with a clinically trained executive.
